---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ispv

<!-- badges: start -->
[![R-CMD-check](https://github.com/petrbouchal/ispv/workflows/R-CMD-check/badge.svg)](https://github.com/petrbouchal/ispv/actions)
<!-- badges: end -->

The goal of ispv is to retrieve and load data from the Czech database of average earnings (ISPV, Informační systém o průměrném výdělku).

## Installation

You can install the development version of ispv like so:

``` r
remotes::install_github("petrbouchal/ispv")
```

## Example

The package does two things, either for national or regional data:

- retrieve links to Excel files on the ISPV site for a given year
- read those data and return in a clean tibble ready for analysis

Here is an example for regional data:

```{r example, message - FALSE, warning=FALSE}
library(ispv)

links <- pv_list_reg(2020)

paths <- file.path(getwd(), links$name[c(1, 2, 15, 16)])
purrr::walk2(links$url[c(1, 2, 15, 16)], paths, curl::curl_download)

data <- pv_reg_monthlypay_isco4(paths)

data
```

```{r, echo=FALSE, include=FALSE}
file.remove(paths)
```


This is a long format data frame, so is easy to analyse and visualise:

```{r example-filter}
library(dplyr)
data_select <- data |> 
  filter(isco4_name == "Mzdoví účetní")
```

```{r example-plot}
library(ggplot2)
ggplot(data_select, aes(pay_mean, kraj_name, fill = sfera)) +
  geom_col() +
  facet_wrap(~sfera, ncol = 1) +
  theme_minimal()
```

## About the data
 
The data is special in several ways:
 
- it contains pay and staff counts broken by ISCO-4 level for each region (kraj), further separated into the "wage sphere" and "salary sphere" - broadly private and public sector, respectively
- pay is broken down into regular pay, bonuses, other compensation and other components
 
This is probably the most detailed labour market data available publicly.

See details on the official page: [English](https://ispv.cz/en/about-ispv.aspx) and [Czech](https://ispv.cz/cz/O-ISPV.aspx)
 
The data is only available to the public in Excel files, separately for each year and region. This package makes it possible to download and read the data in bulk.

The public sector part of the data is based on individual-level returns provided by each public sector organisation. The private sector part is based on a survey.

The data items are documented in the relevant data loading functions, i.e. currently only `pv_load_reg`.
 
## Limitations of the package
 
The source data contains multiple breakdowns and other types of information. 
Currently, the package only supports loading the pay-and-headcount tables for certain breakdowns (notably ISCO and NACE as well as gender and age, and education levels).

## Credits

The data is collected and managed by Trexima on behalf of the Czech Ministry of Labour and Social Affairs.

This package is not affiliated with or endorsed by the ministry, but they deserve credit for commissioning the data collection and maintenance.
